Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-01-21 manuscript.tex
typo

Selective attention is commonly investigated using interference
paradigms such as the Stroop (1935) and flanker (B. A. Eriksen \&
Eriksen, 1974) tasks, where participants identify a target while
ignoring a response-congruent or -incongruent distractor. Performance is
typically better on congruent versus incongruent trials and the
difference---the congruency effect---taken as an index of attentional
priorities. Large congruency effects are thought to reflect ineffective
filtering of the distracting stimuli whereas small congruency effects
are thought to reflect effective filtering. By probing factors that
systematically alter congruency effects, we can then make inferences
about processes that control attentional filtering. For example,
manipulating the frequency of conflict via the proportion of congruent
versus incongruent trials has shown to influence the size of the
congruency effect. Typically, a high proportion congruent experiment
produces large congruency effects, whereas a low proportion congruent
experiment produces small congruency effects (Logan \& Zbrodoff, 1979;
Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually
explained as strategic control, where participants increase attentional
control under high-conflict demands and relax attentional control under
low-conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan,
Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work
however, has demonstrated that attentional control is not only adjusted
by top-down regulation, but can also be triggered automatically by
environmental cues (Brosowsky \& Crump, 2018; J. Bugg \& Crump, 2012;
Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012;
Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (see also, Corballis \& Gratton,
2003; Crump, Gong, \& Milliken, 2006) presented Stroop stimuli in one of
two randomly chosen locations and manipulated the frequency of conflict
associated with each location. One location was associated with a high
frequency of conflict (25\% congruent trials) and the other with a low
frequency of conflict (75\% congruent trials). Overall, the proportion
of congruent trials was 50\% and randomized such that the upcoming
location could not be predicted. Even so, congruency effects were shown
to be smaller for trials where the stimulus appeared in the high
conflict location as compared to the low conflict location. This effect,
known now as the context-specific proportion congruent effect (CSPC),
has now been replicated in a number of different selective attention
paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais,
Harris, Sinanian, \& Bunge, 2015; J. Bugg, 2014; Crump, 2016; Crump,
Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer,
Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

@@

Prior work shows that task-relevance is necessary for establishing and
using context-attention associations. First, the most commonly used
context cue is location (Brosowsky \& Crump, 2016; Corballis \& Gratton,
2003; Crump, 2016; Crump et al., 2006; Crump, Vaquero, \& Milliken,
2008; Weidler \& Bugg, 2016; Weidler et al., 2018). Although in some
sense, the location is irrelevant for identifying the color of a word,
it seems disingenuous to consider it completely task-irrelevant. In
order to identify the target, the participant must first identify and
orient to its location on the screen. In this is sense, it is relevant
to their ability to complete the task and furthermore, location
information may receive priority during encoding (Logan, 1998; Mayr,
1996). Similar arguments could be made about other context features that
are inherent to processing the target stimulus, like font-type (J. Bugg,
Jacoby, \& Toth, 2008) and color (Vietze \& Wendt, 2009).

@@

In light of prior work, we take this as evidence that the contextual
recruitment of selective attention, although likely implicit, is not
obligatory (e.g., Brosowsky \& Crump, 2016), requiring that
environmental information be incorporated into the task representation.
Similarly, there appears to be flexibility in which environmental
features are selected and used to guide attention, which can be rapidly
updated depending on the task-relevance of those features. Such a result
lends some insight into how context-specific control might operate in
more complex, real-world environments, where there is an over-abundance
of environmental features that afford many different learned
associations. From a theoretical perspective, this result is consistent
memory-based accounts of CSPC phenomena (Brosowsky \& Crump, 2018; J.
Bugg \& Hutchison, 2013; Crump \& Milliken, 2009; Crump et al., 2018,
2017). Under this view, a memory process encodes attentional priorities
in the representation of individual experiences and, as a result, become
associated with the environment where they were used. The subsequent
reoccurrence of a prior context triggers the retrieval and reinstatement
of those attentional priorities. Our results show however, that all the
features of the environment may not be treated equally and that only
task-relevant features are used to probe memory and guide attention.

@@

\hypertarget{ref-weidler_attentional_2018}{}
Weidler, B. J., Dey, A., \& Bugg, J. M. (2018). Attentional control
transfers beyond the reference frame. \emph{Psychological Research},
1--14.

>>>>

Selective attention is commonly investigated using interference
paradigms such as the Stroop (1935) and flanker (B. A. Eriksen \&
Eriksen, 1974) tasks, where participants identify a target while
ignoring a response-congruent or -incongruent distractor. Performance is
typically better on congruent versus incongruent trials and the
difference---the congruency effect---taken as an index of attentional
priorities. Large congruency effects are thought to reflect ineffective
filtering of the distracting stimuli whereas small congruency effects
are thought to reflect effective filtering. By probing factors that
systematically alter congruency effects, we can then make inferences
about processes that control attentional filtering. For example,
manipulating the frequency of conflict via the proportion of congruent
versus incongruent trials has shown to influence the size of the
congruency effect. Typically, a high proportion congruent experiment
produces large congruency effects, whereas a low proportion congruent
experiment produces small congruency effects (Logan \& Zbrodoff, 1979;
Lowe \& Mitterer, 1982; West \& Baylis, 1998). This result is usually
explained as strategic control, where participants increase attentional
control under high-conflict demands and relax attentional control under
low-conflict demands (Logan, 1980; Logan \& Zbrodoff, 1979; Logan,
Zbrodoff, \& Williamson, 1984; Lowe \& Mitterer, 1982). Recent work
however, has demonstrated that attentional control is not only adjusted
by top-down regulation, but can also be triggered automatically by
environmental cues (Brosowsky \& Crump, 2018; Bugg \& Crump, 2012;
Egner, 2014; Fischer \& Dreisbach, 2015; King, Korb, \& Egner, 2012;
Mayr \& Bryck, 2007).

For example, Crump, Gong, and Milliken (see also, Corballis \& Gratton,
2003; Crump, Gong, \& Milliken, 2006) presented Stroop stimuli in one of
two randomly chosen locations and manipulated the frequency of conflict
associated with each location. One location was associated with a high
frequency of conflict (25\% congruent trials) and the other with a low
frequency of conflict (75\% congruent trials). Overall, the proportion
of congruent trials was 50\% and randomized such that the upcoming
location could not be predicted. Even so, congruency effects were shown
to be smaller for trials where the stimulus appeared in the high
conflict location as compared to the low conflict location. This effect,
known now as the context-specific proportion congruent effect (CSPC),
has now been replicated in a number of different selective attention
paradigms (e.g., Alards-Tomalin, Brosowsky, \& Mondor, 2017; Blais,
Harris, Sinanian, \& Bunge, 2015; Bugg, 2014; Crump, 2016; Crump,
Milliken, Leboe-McGowan, Leboe-McGowan, \& Gao, 2018; Fischer,
Gottschalk, \& Dreisbach, 2014; Hübner \& Mishra, 2016).

@@

Prior work shows that task-relevance is necessary for establishing and
using context-attention associations. First, the most commonly used
context cue is location (Brosowsky \& Crump, 2016; Corballis \& Gratton,
2003; Crump, 2016; Crump et al., 2006; Crump, Vaquero, \& Milliken,
2008; Weidler \& Bugg, 2016; Weidler et al., 2018). Although in some
sense, the location is irrelevant for identifying the color of a word,
it seems disingenuous to consider it completely task-irrelevant. In
order to identify the target, the participant must first identify and
orient to its location on the screen. In this is sense, it is relevant
to their ability to complete the task and furthermore, location
information may receive priority during encoding (Logan, 1998; Mayr,
1996). Similar arguments could be made about other context features that
are inherent to processing the target stimulus, like font-type (Bugg,
Jacoby, \& Toth, 2008) and color (Vietze \& Wendt, 2009).

@@

In light of prior work, we take this as evidence that the contextual
recruitment of selective attention, although likely implicit, is not
obligatory (e.g., Brosowsky \& Crump, 2016), requiring that
environmental information be incorporated into the task representation.
Similarly, there appears to be flexibility in which environmental
features are selected and used to guide attention, which can be rapidly
updated depending on the task-relevance of those features. Such a result
lends some insight into how context-specific control might operate in
more complex, real-world environments, where there is an over-abundance
of environmental features that afford many different learned
associations. From a theoretical perspective, this result is consistent
memory-based accounts of CSPC phenomena (Brosowsky \& Crump, 2018; Bugg
\& Hutchison, 2013; Crump \& Milliken, 2009; Crump et al., 2018, 2017).
Under this view, a memory process encodes attentional priorities in the
representation of individual experiences and, as a result, become
associated with the environment where they were used. The subsequent
reoccurrence of a prior context triggers the retrieval and reinstatement
of those attentional priorities. Our results show however, that all the
features of the environment may not be treated equally and that only
task-relevant features are used to probe memory and guide attention.

@@

\hypertarget{ref-weidler_attentional_2018}{}
Weidler, B. J., Dey, A., \& Bugg, J. (2018). Attentional control
transfers beyond the reference frame. \emph{Psychological Research},
1--14.
